An Algorithmic Introduction to Lagrangian Coherent Structures

### 2014-01-31 main.tex
Remove graphics filename extension from includegraphics command
Minor modifications to main.tex

Strainlines, trajectories of the $\boldsymbol \xi_1$ vector field, are integrated starting from local maxima of $\lambda_2$ on the main grid. 
The local maximisation distance \lstinline!distance! is defined. This distance is set to a value between two and five grid points, since it represents the number of grid points over which the Cauchy-Green strain tensor is susceptible to numerical noise from integration. Thus, the local maximisation distance is expected to be a function of variable time step integration tolerances. A record is made of all grid points where $\lambda_2$ is maximal within a circle whose radius equals the defined maximisation distance. 
Then the first strainline is integrated from the largest of all the local maxima in both directions, i.e., integrating $\boldsymbol \xi_1$ and $-\boldsymbol \xi_1$. Similarly to lambda-lines, this integration must take into account eigenvector field orientation discontinuities. Prior to initiating strainline integration, a maximum length is defined. Strainlines are then integrated until reaching the domain boundary or the maximum length. The maximum length may be seen as safety factor to prevent highly sinuous strainlines leading to inordinately long computations; the maximum length should be set such that most strainlines stop at the domain boundary. 
Once a strainline position has been calculated, local maxima of $\lambda_2$ within the minimisation distance are flagged as ineligible initial positions for subsequent strainlines. After the first strainline has been obtained and any ineligible local maxima of $\lambda_2$ have been flagged, the next strainline is integrated starting form the highest eligible maximum of $\lambda_2$. The process continues until all eligible local maxima of $\lambda_2$ have been used to initialise strainline integration. 
%Finally, any strainline segment contained within (closed) shear LCS are removed. 
These calculations are performed by the function \lstinline!seed_curves_from_lambda_max!.
Note that once a strainline integration has begun, it may progress so as to be within the local minimisation distance of an existing strainline. In other words, the local minimisation distance controls where strainlines begin, but not where they stop.

@@

Figures~\ref{f:ocean dataset hyperbolic shear lcs details strainline} and \ref{f:ocean dataset hyperbolic shear lcs details stretchline} show elliptic and hyperbolic LCS on 22 November 2006, i.e., the same time instant shown in~\textcite{haller13:_coher_lagran,beron-vera13:_objec_agulh}. Note that here the integration time is $T=30\,\mathrm{days}$, as opposed to $90$\,days in their analysis. Since for $90$\,days the hyperbolic LCS would be a confusing entanglement of lines, we choose a shorter integration time here, to be able to show hyperbolic LCS along with elliptic LCS. 
For the shorter integration time, the analysis with LCS Tool reveals a large coherent eddy bounded by a closed lambda-line at $(3,-32)$ which corresponds to eddy \#2 in Fig.\,3 of \textcite{beron-vera13:_objec_agulh}. Additionally, a further smaller coherent eddy core at $\sim(1,-31)$ is found that does not stay coherent over a time of $90$\, days. Strainlines (red) and stretchlines (blue) mark the core lines of strong normal attraction and repulsion.

%\begin{figure}
%\begin{center}
%\includegraphics[width=\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_strainline}
%\end{center}
%\caption{Ocean dataset lambda-line (green) and strainline (red) LCS.}
%\label{f:ocean dataset hyperbolic shear lcs strainline}
%\end{figure}

%\begin{figure}
%\begin{center}
%\includegraphics[width=\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_stretchline}
%\end{center}
%\caption{Ocean dataset lambda-line (green) and stretchline (blue) LCSs.}
%\label{f:ocean dataset hyperbolic shear lcs stretchline}
%\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.8\textwidth]{graphics/ocean_dataset/hyperbolic_shearline_lcs_details_stretchline.eps}
\end{center}
\caption{Lambda-line LCS (green) and stretchline LCS (blue) of the ocean data set. Background gray-scale is the finite-time Lyapunov exponent. Dashed green lines are Poincare sections. White dots indicate $\lambda_1$ local minima used for stretchline integration initial positions.}
\label{f:ocean dataset hyperbolic shear lcs details stretchline}
\end{figure}

>>>>

Strainlines, trajectories of the $\boldsymbol \xi_1$ vector field, are integrated starting from local maxima of $\lambda_2$ on the main grid. The local maximisation distance \lstinline!distance! is defined. This distance is set to a value between two and five grid points, since it represents the number of grid points over which the Cauchy-Green strain tensor is susceptible to numerical noise from integration. Thus, the local maximisation distance is expected to be a function of variable time step integration tolerances. A record is made of all grid points where $\lambda_2$ is maximal within a circle whose radius equals the defined maximisation distance. Then the first strainline is integrated from the largest of all the local maxima in both directions, i.e., integrating $\boldsymbol \xi_1$ and $-\boldsymbol \xi_1$. Similarly to lambda-lines, this integration must take into account eigenvector field orientation discontinuities. Prior to initiating strainline integration, a maximum length is defined. Strainlines are then integrated until reaching the domain boundary or the maximum length. The maximum length may be seen as safety factor to prevent highly sinuous strainlines leading to inordinately long computations; the maximum length should be set such that most strainlines stop at the domain boundary. Once a strainline position has been calculated, local maxima of $\lambda_2$ within the minimisation distance are flagged as ineligible initial positions for subsequent strainlines. After the first strainline has been obtained and any ineligible local maxima of $\lambda_2$ have been flagged, the next strainline is integrated starting form the highest eligible maximum of $\lambda_2$. The process continues until all eligible local maxima of $\lambda_2$ have been used to initialise strainline integration. These calculations are performed by the function \lstinline!seed_curves_from_lambda_max!. Note that once a strainline integration has begun, it may progress so as to be within the local minimisation distance of an existing strainline. In other words, the local minimisation distance controls where strainlines begin, but not where they stop.

@@

Figures~\ref{f:ocean dataset hyperbolic shear lcs details strainline} and \ref{f:ocean dataset hyperbolic shear lcs details stretchline} show elliptic and hyperbolic LCS on 22 November 2006, i.e., the same time instant shown in~\textcite{haller13:_coher_lagran,beron-vera13:_objec_agulh}. Note that here the integration time is $T=30\,\mathrm{days}$, as opposed to $90$\,days in their analysis. Since for $90$\,days the hyperbolic LCS would be a confusing entanglement of lines, we choose a shorter integration time here, to be able to show hyperbolic LCS along with elliptic LCS. For the shorter integration time, the analysis with LCS Tool reveals a large coherent eddy bounded by a closed lambda-line at $(3,-32)$ which corresponds to eddy \#2 in Fig.\,3 of \textcite{beron-vera13:_objec_agulh}. Additionally, a further smaller coherent eddy core at $\sim(1,-31)$ is found that does not stay coherent over a time of 90 days. Both closed orbits are stable if the resolution for the Cauchy-Green strain tensor field is increased. Strainlines (red) and stretchlines (blue) mark the core lines of strong normal attraction and repulsion.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.8\textwidth]{graphics/ocean_dataset/hyperbolic_shearline_lcs_details_stretchline}
\end{center}
\caption{Lambda-line LCS (green) and stretchline LCS (blue) of the ocean data set. Background gray-scale is the finite-time Lyapunov exponent. Dashed green lines are Poincare sections. White dots indicate $\lambda_1$ local minima used for stretchline integration initial positions.}
\label{f:ocean dataset hyperbolic shear lcs details stretchline}
\end{figure}
